---
title: "Test_object_generation"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script contains code used in generating the example datasets used for demonstration in the manuscript (`AML_Seurat`, `AML_SCE`, and `AML_h5ad()`) from the Triana et al. (2021) AML reference dataset (https://doi.org/10.1038/s41590-021-01059-0).

## Download source data

To download the source data for this script, go to https://figshare.com/articles/dataset/Expression_of_197_surface_markers_and_462_mRNAs_in_15281_cells_from_blood_and_bone_marrow_from_a_young_healthy_donor/13398065/2, and press "Download". Please ensure that "Version 2" is selected. 

## Set up reticulate, Python environment

To create `AML_h5ad()`, the reticulate package must be installed, and a Python environment must be initialized with the following packages:

- Numpy
- Scipy
- Pandas
- AnnData

For more information, see https://github.com/amc-heme/SCUBA?tab=readme-ov-file#pre-requisites.

```{r, warning=FALSE, message=FALSE}
library(Seurat)
library(SingleCellExperiment)
library(reticulate)
library(anndata)
library(HDF5Array)
library(ggplot2)
library(dplyr)
# devtools::install_github("cellgeni/sceasy")
library(sceasy)

# Replace with path to the downloaded dataset
sobj <- 
  readRDS(
    "200AB_projected_v2.rds"
    )
```
 
## Add PCA and UMAP reductions

PCA and UMAP reductions do not exist in this object. To create a format suitable for demonstration in SCUBA, these are created below.

```{r, warning=FALSE, message=FALSE}
# Change default asssay to RNA (so "BOTH" can be removed)
DefaultAssay(sobj) <- "RNA"

# Remove BOTH assay
sobj[["BOTH"]] <- NULL

# Normalization
sobj <-
  NormalizeData(
    object = sobj
    )

# Variable feature selection (will be all features, due to the 
# limited panel)
sobj <-
  FindVariableFeatures(
    sobj
  )

# Scale data
sobj <-
  ScaleData(sobj)

# PCA reduction
sobj <-
  RunPCA(
    sobj
  )

# Clustering, UMAP reduction
sobj <- FindNeighbors(sobj, dims = 1:10)
sobj <- FindClusters(sobj, resolution = 0.4)
sobj <- RunUMAP(sobj, dims = 1:10)
```

## Create lower-resolution cell type variable

The `ct` metadata variable contains high resolution cell type assignments based on the combined gene and antibody modalities. For the sake of simplicity and demonstration, the 38 cell types assigned in `ct` will be condensed to 10 cell types. 

```{r}
sobj@meta.data$condensed_cell_type <-
  case_when(
    sobj@meta.data$ct %in% 
      c("Early erythroid progenitor", 
        "Late erythroid progenitor",
        "Aberrant erythroid",
        "Eosinophil-basophil-mast cell progenitors",
        "Erythro-myeloid progenitors",
        "Megakaryocyte progenitors"
        ) ~ "Erythroid progenitors",
    
    sobj@meta.data$ct %in% 
      c("pro-B cells", 
        "Pre-pro-B cells",
        "Small pre-B cell") ~ "pro-B cells",
    
    sobj@meta.data$ct %in% 
      c("Conventional dendritic cell 1", 
        "Conventional dendritic cell 2", 
        "Dendritic-like blasts") ~ "Dendritic cells",
    
    sobj@meta.data$ct %in% 
      c("Plasmacytoid dendritic cell progenitors", 
        "Plasmacytoid dendritic cells") ~ "Plasmacytoid dendritic cells",
    
    sobj@meta.data$ct %in% 
      c("Lymphomyeloid prog",
        "Early promyelocytes",
        "Late promyelocytes",
        "Pre-B cells",
        "Pro-B cells" 
        ) ~ "Lymphomyeloid Progenitors",
    
    sobj@meta.data$ct %in% 
      c("CD8+ central memory T cells", 
        "CD8+CD103+ tissue resident memory T cells", 
        "CD8+ naive T cells", 
        "CD8+ effector memory T cells",
        "GammaDelta T cells") ~ "CD8+ T Cells",
    
    sobj@meta.data$ct %in% 
      c("Nonswitched memory B cells", 
        "Mature naive B cells", 
        "Class switched memory B cells", 
        "CD11c+ memory B cells", 
        "Immature B cells"
        ) ~ "B Cells",
    
    sobj@meta.data$ct %in% 
      c("CD4+ cytotoxic T cells", 
        "CD4+ memory T cells", 
        "CD4+ naive T cells",
        "CD69+PD-1+ memory CD4+ T cells",
        "Naive CD4+T cells") ~ "CD4+ T Cells",
    
    sobj@meta.data$ct %in% c(
      "CD56dimCD16+ NK cells", 
      "CD16+CD56dim NK cells", 
      "CD56brightCD16- NK cells", 
      "NK T cells") ~ "NK Cells",
    
    # Monocytic cluster
    sobj@meta.data$ct %in% c(
      "Non-classical monocytes",
      "Classical Monocytes",
      "Myelocytes",
      "Monocyte-like blasts" 
    ) ~ "Monocytes",
    
    # Primitive cell cluster
    sobj@meta.data$ct %in% c(
      "Promyelocytes",
      "Early Promyelocytes",
      "Late Promyelocytes",
      "Myeloblasts",
      "Lymphoid-primed multipotent progenitors",
      "NK cell progenitors",
      "HSCs & MPPs",
      "Erythro-myeloid progenitors",
      "Immature-like blasts",
      "Mesenchymal cells_1"
    ) ~ "Primitive",
    
    # If the cell type does not match any of the groups above, use the 
    # current name (value of sobj$ct for the current cell)
    TRUE ~ sobj@meta.data$ct
    )

# Collapse Erythroid and lymphomyeloid progenitors into the primitive cluster
sobj@meta.data$condensed_cell_type <-
  case_when(
    # Primitive cell cluster
    sobj@meta.data$condensed_cell_type %in% c(
      "Lymphomyeloid Progenitors",
      "Erythroid progenitors"
    ) ~ "Primitive",
    
    # If the cell type does not match any of the groups above, use the 
    # current name (value of cell_type for the current cell)
    TRUE ~ sobj@meta.data$condensed_cell_type
    )
```

##### Breaking up of "Monocyte" cluster

The monocyte cluster appears to cluster separately between peripheral bone marrow samples, and blood samples (as shown below, using the `Batch` variable). For demonstration in SCUBA, we split this cluster into PBMC monocytes and bone marrow monocytes. 

```{r}
Seurat::DimPlot(
  sobj, 
  reduction = "umap", 
  group.by = "condensed_cell_type",
  split.by = "Batch"
  ) +
  theme(
    # Size of legend text
      legend.text = element_text(size = 8),
      # Change the spacing between the points (keys) in legend and the text
      legend.key.size = unit(0.5, "points"),
      legend.position = "right"
      ) +
  guides(
      color = 
        guide_legend(
          override.aes = list(size = 2),
          # Number of columns in legend
          ncol = 1
          )
      )
```

```{r}
# Split monocytes cluster based on tissue type (via "Batch")
sobj@meta.data$condensed_cell_type <-
  case_when(
    sobj@meta.data$condensed_cell_type == "Monocytes" & 
      sobj@meta.data$Batch == "BM_200AB" ~ "BM Monocytes",
    sobj@meta.data$condensed_cell_type == "Monocytes" & 
      sobj@meta.data$Batch == "PBMC_200AB" ~ "PBMC Monocytes",
    TRUE ~ sobj@meta.data$condensed_cell_type
  )

# Visualize result
Seurat::DimPlot(
  sobj, 
  reduction = "umap", 
  group.by = "condensed_cell_type",
  split.by = "Batch"
  ) +
  theme(
    # Size of legend text
    legend.text = element_text(size = 8),
    # Change the spacing between the points (keys) in legend and the text
    legend.key.size = unit(0.5, "points"),
    legend.position = "right"
    ) +
  guides(
    color = 
      guide_legend(
        override.aes = list(size = 2),
        # Number of columns in legend
        ncol = 1
        )
    )
```

## Downsampling

At this point, the object has sufficient reductions and cell type metadata at a resolution manageable for demonstration. Additionally, the objects used for demonstrated in the SCUBA manuscript were slimmed down to a size of several megabytes, so they could be included directly in the SCUBA repo as test datasets. This was done below via downsampling.

```{r}
set.seed(325)

# Subset while downsampling on condensed_cell_type
AML_Seurat <- 
  subset(
    SetIdent(sobj, value = "condensed_cell_type"),
    downsample = 25
    ) 
```

## Generation of SingleCellExperiment, anndata objects

To demonstrate the input flexibility of SCUBA, the downsampled dataset was converted to anndata and SingleCellExperiment formats.

```{r}
# Convert object to SCE format
AML_sce <- as.SingleCellExperiment(AML_Seurat, assay = "RNA")
AML_sce_ab <- as.SingleCellExperiment(AML_Seurat, assay = "AB")

altExps(AML_sce) <- list("AB" = AML_sce_ab)

# The code below was used to save the object using HDF5 storage. While the
# benefits of this are not realized with a small dataset like this one, 
# RAM consumption can be considerably decreased for large datasets. The usage
# of SCUBA does not differ 
# saveHDF5SummarizedExperiment(
#   AML_sce, 
#   dir = "AML_sce"
#   )
# 
# AML_sce <- loadHDF5SummarizedExperiment(
#   dir = "AML_sce"
# )
```

```{r}
# Convert RNA modality first, then add AB modality
AML_h5ad <- 
  sceasy::convertFormat(
    AML_Seurat,
    from = "seurat",
    to = "anndata",
    main_layer = "data",
    drop_single_values = FALSE
    )

AB_modality <-
    sceasy::convertFormat(
      AML_Seurat,
      from = "seurat",
      to = "anndata",
      main_layer = "data",
      assay = "AB",
      drop_single_values = FALSE
    )

# Append the object created for the AB modality as a matrix
# By convention, additional modalities are stored in the obsm slot.
# Modalities must be stored in the obsm slot to be accessed by SCUBA.
# AB/ADT modalities are also named "protein" by convention, but this is 
# not required for SCUBA access.
AML_h5ad$obsm[["protein"]] <- AB_modality$to_df()
```

