---
title: "SCUBA Manuscript Performance Plots"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

The Rmd below contains code for generating **Figure 4** in the SCUBA manuscript. The dataset for this exercise was downloaded from CELLxGENE. The dataset may be directly downloaded via this link: https://datasets.cellxgene.cziscience.com/9343817c-5c74-4548-a7d5-01990df5af7a.rds

The directory `working_dir` specified below must be set to the directory where the dataset was downloaded for this Rmd to work as-is. Several directories will be created in this directory for storing randomized subsets created from the original dataset.

The randomized subsets created by this Rmd will take up large amounts of disk space. Please allow approximately 175 GB of free disk space to run this Rmd. 

The disk space requirements are related to the approach taken for performance benchmarking. Initially the object was downloaded and randomized subsets of each object class were created in-memory, but large decreases in performance were observed for objects with large numbers of cells. This was likely due to the usage of swap (or the paging file on Windows), as the size of the object and the current randomized subset exceeded the amount of physical memory available. In order for performance testing to be performed on a personal computer in a way that excludes the confounding factor of swap usage, we decided to store subsets on disk instead of creating them in memory. 

```{r}
library(SCUBA)
# BiocManager::install("HDF5Array")
library(HDF5Array)
library(Seurat)
# BiocManager::install("SingleCellExperiment")
library(SingleCellExperiment)
library(anndata)
library(reticulate)

library(dplyr)
library(ggplot2)

library(tictoc)
library(cowplot)
library(stringr)

# Used to add gene symbols to objects downloaded from CELLxGENE
# Feature names of CELLxGENE objects are ENSEMBL IDs by default
# BiocManager::install("biomaRt")
library(biomaRt)
#devtools::install_github("cellgeni/sceasy")
library(sceasy)

# Use cowplot ggplot2 theme by default
theme_set(theme_cowplot())

# Functions 
# Removes grey background from facet labels, and bolds feature names
facet_theme <- 
  theme(
    strip.background = element_rect(fill = "#FFFFFF00"), 
    strip.text = element_text(face = "bold")
    )

# Path to a directory where files and directories for this analysis will be stored.
working_dir <- "working_dir_path"
```

```{r}
# Load Seurat object
sobj <- 
    readRDS(
      paste0(
        working_dir,
        "9343817c-5c74-4548-a7d5-01990df5af7a.rds"
      )
    )
```

# Create Subsets and store in memory 

Subsets for Seurat objects

```{r}
# Set subset sizes to create, and number of randomized subsets per size
subset_sizes <- c(1000, 2000, 5000, 10000, 20000, 50000, 100000, 200000, 500000) |> as.integer()
trials <- 1:5
```


```{r}
# Form random subsets for testing
for (subset_size in subset_sizes){
    for (i in trials){
        cat(subset_size, " cells, ", "trial ", i, sep = "")
        
        # Random subsetting
        # Take a random sample of n cells per trial 
        # (n will increment, to test scalability of performance)
        subset <- 
            sobj[, sample(x = 1:ncol(sobj), size = subset_size, replace = FALSE)]
        
        # Form filename for subset
        subset_filename <- paste0(subset_size, "_", i, ".rds")
        
        # Save subset to disk
        saveRDS(
            subset, 
            file = 
                paste0(
                  working_dir,
                  "Random_subsets/Seurat/",
                  subset_filename
                )
            )
        cat("\r")
    }
}

```

SingleCellExperiment objects

```{r}
# Generate SCE object from each .Rds object from above code block
for (subset_size in subset_sizes){
    for (i in trials){ 
        gc()
        rds_path <- 
            paste0(
              working_dir,
              "Random_subsets/Seurat/",
              # as.integer to output 500000 instead of 5e+05
              paste0(as.integer(subset_size), "_", i, ".rds")
            )
        
        cat(
            subset_size, 
            " cells, ", 
            "trial ", 
            i, 
            " load Seurat object ", 
            rds_path, 
            sep = ""
            )
         cat("\r")
        
        # Load Seurat object for the ith trial of the current subset size,
        # and convert to an SCE object
        sobj <- 
            readRDS(
                rds_path
            )
        
        cat(subset_size, " cells, ", "trial ", i, " make SCE object", sep = "")
        cat("\r")
        
        # Only one assay present
        sce <- as.SingleCellExperiment(sobj, assay = "RNA")
        
        sce_path <- 
            paste0(
              working_dir,
              "Random_subsets/SingleCellExperiment/",
              paste0(as.integer(subset_size), "_", i)
            )
        
        cat(
            subset_size, 
            " cells, ", 
            "trial ", 
            i, 
            " save SCE object ", 
            sce_path, 
            sep = ""
            )
    
        # Save in HDF5 format (use subset size and trial number)
        saveHDF5SummarizedExperiment(
          sce, 
          dir = sce_path,
          level = 3,
          replace = TRUE
          )
        
        cat("\r")
        }
    }



```

anndata objects 

```{r}
#subset_sizes <- c(1000, 2000, 5000, 10000, 20000, 50000, 100000, 200000, 500000) |> as.integer()
#subset_sizes <- 500000
#trials <- 1:5

# Generate anndata object from each .Rds object from above code block
for (subset_size in subset_sizes){
    for (i in trials){ 
        rds_path <- 
            paste0(
              working_dir,
              "Random_subsets/Seurat/",
              # as.integer to output 500000 instead of 5e+05
              paste0(as.integer(subset_size), "_", i, ".rds")
            )
        
        cat(
            subset_size, 
            " cells, ", 
            "trial ", 
            i, 
            " load Seurat object ", 
            rds_path, 
            sep = ""
            )
         cat("\r")
        
        # Load Seurat object for the ith trial of the current subset size,
        # and convert to an anndata object
        sobj <-
          readRDS(
            rds_path
            )
        
        adata_path <- 
          paste0(
            working_dir,
            "Random_subsets/anndata/",
            paste0(as.integer(subset_size), "_", i, ".h5ad")
            )
        
        cat(
          subset_size, 
          " cells, ", 
          "trial ", 
          i, 
          " save Anndata object ", 
          adata_path, 
          sep = ""
          )
    
        # Convert to anndata object and save
        adata <- 
          sceasy::convertFormat(
            sobj, 
            from = "seurat", 
            to = "anndata",
            drop_single_values = FALSE
            )
        
        anndata::write_h5ad(
          adata, 
          filename = adata_path
          )
        
        cat("\r")
        }
    }




```

# Performance testing of FetchData

The testing below replicates the results shown in **Figure 4C** of the manuscript. Data for one feature was retrieved from each of the subsets using FetchData, and the run time was measured via the `tictoc` package. The performance of the existing Seurat method was compared to the methods created by SCUBA for SingleCellExperiment and anndata objects.

```{r}
times <- c()
n_cells <- c()

# Seurat objects
for (subset_size in subset_sizes){
    # Run test for i trials per number of cells
    for (i in trials){
        # Display the current subset size and trial
        cat("Subset size ", subset_size, ", trial ", i, sep = "")
        
        # Load random subset
        subset <- 
          readRDS(
            paste0(
              working_dir,
              "Random_subsets/Seurat/",
              paste0(as.integer(subset_size), "_", i, ".rds")
            )
          )
        
        # Run FetchData and time run with Sys.time()
        start_time <- Sys.time()
        
        FetchData(
            subset, 
            # CellXGene displays feature names according to Ensembl ID.
            # To avoid the hassle of transforming cell IDs to gene symbols, 
            # I put a random gene symbol in GeneCards and used the corresponding
            # ensembl ID as the target var.
            vars = c("ENSG00000078399")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        # Also store the number of cells associated with the current trial
        n_cells <- c(n_cells, subset_size)
        # Carriage return, to avoid cluttering output during testing
        cat("\r")
    }
}

seurat_trial <-
    data.frame(
    `n_cells` = n_cells,
    `run_time` = times
    )

seurat_trial$fxn <- "FetchData.Seurat"
```

```{r}
# FetchData, SingleCellExperiment
times <- c()
n_cells <- c()

for (subset_size in subset_sizes){
    # Run test for i trials per number of cells
    for (i in trials){
        # Display the current subset size and trial
        cat("Subset size ", subset_size, ", trial ", i, sep = "")
        
        # Load random subset
        subset <- 
          HDF5Array::loadHDF5SummarizedExperiment(
            paste0(
              working_dir,
              "Random_subsets/SingleCellExperiment/",
              paste0(as.integer(subset_size), "_", i)
              )
            )
        
        # Run FetchData and time run with Sys.time()
        start_time <- Sys.time()
        
        FetchData(
            subset, 
            vars = c("ENSG00000078399")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        # Also store the number of cells associated with the current trial
        n_cells <- c(n_cells, subset_size)
        # Carriage return, to avoid cluttering output during testing
        cat("\r")
    }
}

sce_trial <-
  data.frame(
    `n_cells` = n_cells,
    `run_time` = times
  )

sce_trial$fxn <- "FetchData.SingleCellExperiment"
```

```{r}
# FetchData, anndata objects
times <- c()
n_cells <- c()

for (subset_size in subset_sizes){
    # Run test for i trials per number of cells
    for (i in trials){
        # Display the current subset size and trial
        cat("Subset size ", subset_size, ", trial ", i, sep = "")
        
        # Load random subset
        subset <- 
          anndata::read_h5ad(
            paste0(
              working_dir,
              "Random_subsets/anndata/",
              paste0(as.integer(subset_size), "_", i, ".h5ad")
            )
          )
        
        # Run FetchData and time run with Sys.time()
        start_time <- Sys.time()
        
        FetchData(
            subset, 
            vars = c("ENSG00000078399")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        # Also store the number of cells associated with the current trial
        n_cells <- c(n_cells, subset_size)
        # Carriage return, to avoid cluttering output during testing
        cat("\r")
        
        rm(subset)
        gc()
    }
}

anndata_trial <-
    data.frame(
    `n_cells` = n_cells,
    `run_time` = times
    )

anndata_trial$fxn <- "FetchData.AnnDataR6"

```

```{r}
# Combine results for each object and save
fetchdata_results <- rbind(seurat_trial, sce_trial, anndata_trial)

results_dir <- paste0(working_dir, "performance_results/")

if(!dir.exists(results_dir)) dir.create(results_dir)

write.csv(
  fetchdata_results, 
  file = paste0(working_dir, "performance_results/FetchData.csv"), 
  row.names = FALSE
)
```

# fetchData on one metadata variable vs. fetch_metadata

The testing below replicates the results shown in **Figure 4A** of the manuscript. The script compares the run time of fetching one metadata feature using FetchData compared to using the `fetch_metadata` convenience function provided by SCUBA.

```{r}
#subset_sizes <- c(1000, 2000, 5000, 10000, 20000, 50000, 100000, 200000)
#trials <- 1:5

times <- c()
n_cells <- c()
fxn <- c()

# Seurat objects
for (subset_size in subset_sizes){
    # Run test for i trials per number of cells
    for (i in trials){
        # Display the current subset size and trial
        cat("Subset size ", subset_size, ", trial ", i, sep = "")
        
        # Load random subset
        subset <- 
          readRDS(
            paste0(
              working_dir, 
              "Random_subsets/Seurat/",
              paste0(as.integer(subset_size), "_", i, ".rds")
            )
          )
        
        # Run FetchData and time run with Sys.time()
        start_time <- Sys.time()
        
        FetchData(
            subset, 
            vars = c("cell_type")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        # Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "FetchData.Seurat")
        
        start_time <- Sys.time()
        
        SCUBA::fetch_metadata(
            subset, 
            vars = c("cell_type")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        # Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "fetch_metadata.Seurat")
        
        # Carriage return, to avoid cluttering output during testing
        cat("\r")
    }
}

seurat_trial <-
    data.frame(
    `n_cells` = n_cells,
    `run_time` = times,
    `fxn` = fxn
    )
```

```{r}
# FetchData, SingleCellExperiment
#subset_sizes <- c(1000, 2000, 5000, 10000, 20000, 50000, 100000, 200000)
#trials <- 1:5

times <- c()
n_cells <- c()
fxn <- c()

for (subset_size in subset_sizes){
    # Run test for i trials per number of cells
    for (i in trials){
        # Display the current subset size and trial
        cat("Subset size ", subset_size, ", trial ", i, sep = "")
        
        # Load random subset
        subset <- 
          HDF5Array::loadHDF5SummarizedExperiment(
            paste0(
              working_dir,
              "Random_subsets/SingleCellExperiment/",
              paste0(as.integer(subset_size), "_", i)
              )
            )
        
        # Run FetchData and time run with Sys.time()
        start_time <- Sys.time()
        
        FetchData(
            subset, 
            vars = c("cell_type")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        #  Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "FetchData.SingleCellExperiment")
        
        start_time <- Sys.time()
        
        SCUBA::fetch_metadata(
            subset, 
            vars = c("cell_type")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        # Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "fetch_metadata.SingleCellExperiment")
        # Carriage return, to avoid cluttering output during testing
        cat("\r")
    }
}

sce_trial <-
  data.frame(
    `n_cells` = n_cells,
    `run_time` = times,
    `fxn` = fxn
  )
```

```{r}
#subset_sizes <- c(1000, 2000, 5000, 10000, 20000, 50000, 100000, 200000)
#trials <- 1:5

# FetchData, anndata objects
times <- c()
n_cells <- c()
fxn <- c()

# Seurat objects
for (subset_size in subset_sizes){
    # Run test for i trials per number of cells
    for (i in trials){
        # Display the current subset size and trial
        cat("Subset size ", subset_size, ", trial ", i, sep = "")
        
        # Load random subset
        subset <- 
          anndata::read_h5ad(
            paste0(
              working_dir,
              "Random_subsets/anndata/",
              paste0(as.integer(subset_size), "_", i, ".h5ad")
            )
          )
        
        # Run FetchData and time run with Sys.time()
        start_time <- Sys.time()
        
        FetchData(
            subset, 
            vars = c("cell_type")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        #  Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "FetchData.AnnDataR6")
        
        start_time <- Sys.time()
        
        SCUBA::fetch_metadata(
            subset, 
            vars = c("cell_type")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        
        # Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "fetch_metadata.AnnDataR6")
        
        # Carriage return, to avoid cluttering output during testing
        cat("\r")
    }
}

anndata_trial <-
    data.frame(
    `n_cells` = n_cells,
    `run_time` = times,
    `fxn` = fxn
    )

#have to run this after big anndata chunks
reticulate::import("gc")$collect()

```

```{r}
fetchdata_vs_fetchmetadata_results <- 
  rbind(
    seurat_trial, 
    sce_trial, 
    anndata_trial
  )

write.csv(
  fetchdata_vs_fetchmetadata_results,
  file = paste0(
    working_dir,
    "performance_results/FetchData_vs_fetch_metadata_3.csv"
  ),
  row.names = FALSE
)
```

# fetchData on UMAP coordinates vs. fetch_reduction

The testing below replicates the results shown in **Figure 4B** of the manuscript. The script compares the run time of fetching reduction coordinates for one dimensional reduction using FetchData compared to using the `fetch_reduction` convenience function provided by SCUBA.

```{r}
#subset_sizes <- c(1000, 2000, 5000, 10000, 20000, 50000, 100000, 200000)
#trials <- 1:5

times <- c()
n_cells <- c()
fxn <- c()

# Seurat objects
for (subset_size in subset_sizes){
    # Run test for i trials per number of cells
    for (i in trials){
        # Display the current subset size and trial
        cat("Subset size ", subset_size, ", trial ", i, sep = "")
        
        # Load random subset
        subset <- 
          readRDS(
            paste0(
              working_dir,
              "Random_subsets/Seurat/",
              paste0(as.integer(subset_size), "_", i, ".rds")
            )
          )
        
        # Run FetchData and time run with Sys.time()
        start_time <- Sys.time()
        
        FetchData(
            subset, 
            vars = c("UMAP_1", "UMAP_2")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        # Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "FetchData.Seurat")
        
        start_time <- Sys.time()
        
        SCUBA::fetch_reduction(
            subset, 
            reduction = "UMAP",
            dims = c(1, 2)
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        # Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "fetch_reduction.Seurat")
        
        # Carriage return, to avoid cluttering output during testing
        cat("\r")
    }
}

seurat_trial <-
    data.frame(
    `n_cells` = n_cells,
    `run_time` = times,
    `fxn` = fxn
    )
```

```{r}
# FetchData, SingleCellExperiment
#subset_sizes <- c(1000, 2000, 5000, 10000, 20000, 50000, 100000, 200000)
#trials <- 1:5

times <- c()
n_cells <- c()
fxn <- c()

for (subset_size in subset_sizes){
    # Run test for i trials per number of cells
    for (i in trials){
        # Display the current subset size and trial
        cat("Subset size ", subset_size, ", trial ", i, sep = "")
        
        # Load random subset
        subset <- 
          HDF5Array::loadHDF5SummarizedExperiment(
            paste0(
              working_dir,
              "Random_subsets/SingleCellExperiment/",
              paste0(as.integer(subset_size), "_", i)
              )
            )
        
        # Run FetchData and time run with Sys.time()
        start_time <- Sys.time()
        
        FetchData(
            subset, 
            vars = c("UMAP_1", "UMAP_2")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        #  Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "FetchData.SingleCellExperiment")
        
        start_time <- Sys.time()
        
        SCUBA::fetch_reduction( 
            subset, 
            reduction = "UMAP",
            dims = c(1, 2)
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        # Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "fetch_reduction.SingleCellExperiment")
        # Carriage return, to avoid cluttering output during testing
        cat("\r")
    }
}

sce_trial <-
  data.frame(
    `n_cells` = n_cells,
    `run_time` = times,
    `fxn` = fxn
  )
```

```{r}
#subset_sizes <- c(1000, 2000, 5000, 10000, 20000, 50000, 100000, 200000)
#trials <- 1:5

# FetchData, anndata objects
times <- c()
n_cells <- c()
fxn <- c()

# Seurat objects
for (subset_size in subset_sizes){
    # Run test for i trials per number of cells
    for (i in trials){
        # Display the current subset size and trial
        cat("Subset size ", subset_size, ", trial ", i, sep = "")
        
        # Load random subset
        subset <- 
          anndata::read_h5ad(
            paste0(
              working_dir,
              "Random_subsets/anndata/",
              paste0(as.integer(subset_size), "_", i, ".h5ad")
            )
          )
        
        # Run FetchData and time run with Sys.time()
        start_time <- Sys.time()
        
        FetchData(
            subset, 
            vars = c("X_umap_1", "X_umap_2")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        #  Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "FetchData.AnnDataR6")
        
        start_time <- Sys.time()
        
        SCUBA::fetch_reduction(
            subset, 
            reduction = "X_umap",
            dims = c(1, 2)
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        
        # Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "fetch_reduction.AnnDataR6")
        
        # Carriage return, to avoid cluttering output during testing
        cat("\r")
    }
}

anndata_trial <-
    data.frame(
    `n_cells` = n_cells,
    `run_time` = times,
    `fxn` = fxn
    )

#required after large anndata chunks
reticulate::import("gc")$collect()
```

```{r}
fetchdata_vs_fetchreduction_results <- 
  rbind(seurat_trial, sce_trial, anndata_trial)

write.csv(
   fetchdata_vs_fetchreduction_results,
   file = paste0(
     working_dir,
     "performance_results/FetchData_vs_fetch_reduction.csv"
   ),
   row.names = FALSE
)
```

# Plot results of performance testing 

Once the CSVs in above sections are created, only this section of the Rmd and the following one ("Combined Figure") needs to be ran. 

```{r}
# Define Paths to each results CSV
fetchData_path <- 
  paste0(
     working_dir,
     "performance_results/FetchData.csv"
   )

fetchdata_vs_fetchmetadata_path <-
  paste0(
     working_dir,
     "performance_results/FetchData_vs_fetch_metadata_3.csv"
   )

fetchdata_vs_fetchreduction_path <-
  paste0(
     working_dir,
     "performance_results/FetchData_vs_fetch_reduction.csv"
   )

# Palette used for SCUBA manuscript, fig1: Okabe-Ito palette, rearranged
scuba_manuscript_pal <- palette.colors(palette = "Okabe-Ito")
scuba_manuscript_pal <- scuba_manuscript_pal[c(6,7,4,2,3,8,1,5,9)]
#show_col(scuba_manuscript_pal)
```

```{r fetchData plot}
results_df <- read.csv(fetchData_path)

# Rename function column lavels
results_df$fxn <- 
  plyr::revalue(
    results_df$fxn, 
    c("FetchData.Seurat" = "Seurat", 
      "FetchData.SingleCellExperiment" = "SingleCellExperiment", 
      "FetchData.AnnDataR6" = "anndata"
      )
    )

# Re-factor
results_df$fxn <- 
  factor(
    results_df$fxn, 
    levels = 
      c("Seurat",
        "SingleCellExperiment",
        "anndata"
        )
    )

# Convert number of cells to character
results_df$n_cells <-
  as.character(
    # Extract from scientific notation with as.integer
    as.integer(results_df$n_cells)
    )

results_df$n_cells <- 
  factor(
    results_df$n_cells,
    levels = 
      as.character(
        str_sort(
          unique(results_df$n_cells), 
          numeric = TRUE
          )
        )
  )

results_df <- results_df %>% rename(c("Method" = "fxn"))

fetchData_comparison <- 
  ggplot(data = results_df, aes(x = n_cells, y = run_time)) +
  geom_boxplot(aes(color = Method), linewidth = 0.25, outlier.size = 0.25) +
  scale_color_manual(
    name = NULL,
    values = scuba_manuscript_pal[c(9,1,3)],
    aesthetics = "color",
    na.value = "grey50"
    ) +
  # labs(
  #   title = "FetchData with a Single Gene", 
  #   #subtitle = "Assessed over 100 trials, on full AML dataset, with one feature"
  #   ) +
  xlab("Number of cells") +
  ylab("Run Time (s)") +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    plot.subtitle = element_text(hjust = 0.5),
    #axis.title.y = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)#,
    #legend.position = "bottom"
  ) #+
  #coord_flip()

fetchData_comparison
```

```{r fetchData_vs_fetch_metadata_boxplot}
results_df <- read.csv(fetchdata_vs_fetchmetadata_path)

# Add information for object type and function based on `fxn` column
results_df$object <- 
  case_when(
    results_df$fxn %in% 
      c("FetchData.Seurat", "fetch_metadata.Seurat") ~ "Seurat",
    results_df$fxn %in% 
      c("FetchData.SingleCellExperiment", 
        "fetch_metadata.SingleCellExperiment") ~ "SingleCellExperiment",
    results_df$fxn %in%
      c("FetchData.AnnDataR6", "fetch_metadata.AnnDataR6") ~ "anndata",
    TRUE ~ NA
  )

results_df$Function <- 
  case_when(
    results_df$fxn %in% 
      c("FetchData.Seurat", 
        "FetchData.SingleCellExperiment", 
        "FetchData.AnnDataR6") ~ "FetchData",
    results_df$fxn %in% 
      c("fetch_metadata.Seurat",  
        "fetch_metadata.SingleCellExperiment",
        "fetch_metadata.AnnDataR6") ~ "fetch_metadata",
    TRUE ~ NA
  )

# Re-factor object, function columns
results_df$object <-
  factor(
    results_df$object,
    levels = c("Seurat", "SingleCellExperiment", "anndata")
  )

results_df$Function <-
  factor(
    results_df$Function,
    levels = c("FetchData", "fetch_metadata")
  )


# Convert number of cells to character
results_df$n_cells <-
  as.character(
    # Extract from scientific notation with as.integer
    as.integer(results_df$n_cells)
    )

results_df$n_cells <- 
  factor(
    results_df$n_cells,
    levels = 
      str_sort(
        unique(results_df$n_cells), 
        numeric = TRUE
        )
  )

# Remove single outlier for SingleCellExperiment, FetchData, 10,000 cells
#results_df <- results_df[-113,] 

# z-score results *by test and number of cells* and remove 
# extreme outliers (>= 3 SD)
# Create table with mean, SD for each fxn/n_cells combo
scaling_df <- 
  results_df %>% 
    group_by(fxn, n_cells) %>% 
    summarise(
      mean = mean(run_time, na.rm = TRUE),
      SD = sd(run_time, na.rm = TRUE)
      )

results_df <-
  results_df %>% 
  left_join(
    .,
    scaling_df,
    by = c("fxn", "n_cells")
  ) %>% 
  dplyr::mutate(run_time_scaled = (run_time - mean)/SD)

# Plot
fetchData_vs_fetch_metadata <-
  results_df %>%
  ggplot(data = ., aes(x = n_cells, y = run_time)) +
  geom_boxplot(aes(color = Function), linewidth = 0.25, outlier.size = 0.25) +
  facet_wrap(
    ~ object,
    scales = "free",
    ncol = 1
    ) +
  scale_color_manual(
    name = NULL,
    values = scuba_manuscript_pal[c(7,5)], # #yellow "#F2AF0D", #orange "#F25E0D", #green "#59A662", #Red "#E61A1A"
    aesthetics = "color",
    na.value = "grey50"
    ) +
  # labs(
  #   title = "FetchData vs. fetch_metadata", 
  #   #subtitle = "Assessed over 100 trials, on full AML dataset, with one feature"
  #   ) +
  ylab("Run Time (s)") +
  xlab("Number of Cells") +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    plot.subtitle = element_text(hjust = 0.5),
    #axis.title.y = element_blank(),
    #axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  #coord_cartesian(ylim = c(0,0.5)) +
  facet_theme
 
fetchData_vs_fetch_metadata
```

```{r FetchData vs. fetch_reduction boxplot}
results_df <- 
  read.csv(fetchdata_vs_fetchreduction_path)

# Add information for object type and function based on `fxn` column
results_df$object <- 
  case_when(
    results_df$fxn %in% 
      c("FetchData.Seurat", "fetch_reduction.Seurat") ~ "Seurat",
    results_df$fxn %in% 
      c("FetchData.SingleCellExperiment", 
        "fetch_reduction.SingleCellExperiment") ~ "SingleCellExperiment",
    results_df$fxn %in%
      c("FetchData.AnnDataR6", "fetch_reduction.AnnDataR6") ~ "anndata",
    TRUE ~ NA
  )

results_df$Function <- 
  case_when(
    results_df$fxn %in% 
      c("FetchData.Seurat", 
        "FetchData.SingleCellExperiment", 
        "FetchData.AnnDataR6") ~ "FetchData",
    results_df$fxn %in% 
      c("fetch_reduction.Seurat",  
        "fetch_reduction.SingleCellExperiment",
        "fetch_reduction.AnnDataR6") ~ "fetch_reduction",
    TRUE ~ NA
  )

# Re-factor object, function columns
results_df$object <-
  factor(
    results_df$object,
    levels = c("Seurat", "SingleCellExperiment", "anndata")
  )

results_df$Function <-
  factor(
    results_df$Function,
    levels = c("FetchData", "fetch_reduction")
  )


# Convert number of cells to character
results_df$n_cells <-
  as.character(
    # Extract from scientific notation with as.integer
    as.integer(results_df$n_cells)
    )

results_df$n_cells <- 
  factor(
    results_df$n_cells,
    levels = 
      str_sort(
        unique(results_df$n_cells), 
        numeric = TRUE
        )
  )

# Remove extreme outlier for SingleCellExperiment, fetch_reduction
#results_df <- results_df[-93,]

# z-score results *by test and number of cells* and remove 
# extreme outliers (>= 3 SD)
# Create table with mean, SD for each fxn/n_cells combo
scaling_df <- 
  results_df %>% 
    group_by(fxn, n_cells) %>% 
    summarise(
      mean = mean(run_time, na.rm = TRUE),
      SD = sd(run_time, na.rm = TRUE)
      )

results_df <-
  results_df %>% 
  left_join(
    .,
    scaling_df,
    by = c("fxn", "n_cells")
  ) %>% 
  dplyr::mutate(run_time_scaled = (run_time - mean)/SD)


fetchdata_vs_fetch_reduction <-
  results_df %>%
  ggplot(data = ., aes(x = n_cells, y = run_time)) +
  geom_boxplot(aes(color = Function), linewidth = 0.25, outlier.size = 0.25) +
  facet_wrap(
    ~ object,
    scales = "free",
    ncol = 1
    ) +
  scale_color_manual(
    name = NULL,
    values = scuba_manuscript_pal[c(7,2)],
      #c("#0079FF", "#E61A1A"), # #yellow "#F2AF0D", #orange "#F25E0D", #green "#59A662", #Red "#E61A1A"
    aesthetics = "color",
    na.value = "grey50"
    ) +
  # labs(
  #   title = "FetchData vs. fetch_metadata", 
  #   #subtitle = "Assessed over 100 trials, on full AML dataset, with one feature"
  #   ) +
  ylab("Run Time (s)") +
  xlab("Number of Cells") +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    plot.subtitle = element_text(hjust = 0.5),
    #axis.title.y = element_blank(),
    #axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  #coord_cartesian(ylim = c(0,0.5)) +
  facet_theme
 
fetchdata_vs_fetch_reduction
```

# Combined figure (Figure 4 in manuscript)

```{r}
layout <- 
  c("
    AABB
    AABB
    CCCC
    ")

combined_figure <-
  patchwork::wrap_plots(
    fetchData_vs_fetch_metadata, #+ 
      #theme(legend.position = "bottom"),
    fetchdata_vs_fetch_reduction, #+
      #theme(legend.position = "bottom"),
    fetchData_comparison,
    design = layout,
    guides = "keep"
    ) +
    patchwork::plot_annotation(
        tag_levels = list(c("A", "B", "C"))
      ) &
  theme(plot.tag = element_text(size = 12, face = "bold"))

combined_figure <-
  combined_figure &
    theme(
      legend.text = element_text(size = 8),
      legend.justification = "center",
      axis.title = element_text(size = 9),
      axis.text = element_text(size = 8),
      strip.text = element_text(size = 9),
      legend.position = "bottom"
    )

mm_to_in <- 
  function(mm){
    mm/25.4
  }

pdf(
  file = 
    paste0(
      working_dir,
      "performance_results/fig4.pdf"
      ), 
  width = mm_to_in(150), 
  height = mm_to_in(275),
  pointsize = 8
  )
combined_figure
dev.off()
```
