---
title: "SCUBA Manuscript Performance Plots"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

This Rmd contains the code used to generate the performance plots for the SCUBA manuscript.

```{r}
library(SCUBA)
library(HDF5Array)
library(Seurat)
library(SingleCellExperiment)
library(anndata)
library(reticulate)

library(dplyr)
library(ggplot2)

library(tictoc)
library(cowplot)

library(stringr)

# Used to add gene symbols to objects downloaded from CELLxGENE
# Feature names of CELLxGENE objects are ENSEMBL IDs by default
library(biomaRt)

library(sceasy)

# Use cowplot ggplot2 theme by default
theme_set(theme_cowplot())

# Functions 
# Removes grey background from facet labels, and bolds feature names
facet_theme <- 
  theme(
    strip.background = element_rect(fill = "#FFFFFF00"), 
    strip.text = element_text(face = "bold")
    )

# Path to a directory where files and directories for this analysis will be 
# stored. 
working_dir <- "working_dir_path"
```

For this exercise, I downloaded a larger dataset from CellXGene (the Human Brain atlas dataset):
https://cellxgene.cziscience.com/collections/283d65eb-dd53-496d-adb7-7570c7caa443

direct link: https://datasets.cellxgene.cziscience.com/9343817c-5c74-4548-a7d5-01990df5af7a.rds

I used "all non-neuronal cells", and downloaded the dataset in Seruat format.

```{r}
# Load Seurat object
sobj <- 
    readRDS(
      paste0(
        working_dir,
        "9343817c-5c74-4548-a7d5-01990df5af7a.rds"
      )
    )
```

Due to the very large size of this object, performance issues likely related to swap usage were observed during testing, which introduces a confounding factor when measuring performance. To control this, subsets were formed before testing and saved to disk, instead of being formed during testing with the full object still loaded. 

The original code used for testing is at the bottom of this document, under the `Original Performance Testing Code` section.

# Create Subsets 

Subsets for Seurat objects

```{r}
# Set subset sizes to create, and number of randomized subsets per size
subset_sizes <- c(1000, 2000, 5000, 10000, 20000, 50000, 100000, 200000, 500000) |> as.integer()
trials <- 1:5
```


```{r}
# Form random subsets for testing
for (subset_size in subset_sizes){
    for (i in trials){
        cat(subset_size, " cells, ", "trial ", i, sep = "")
        
        # Random subsetting
        # Take a random sample of n cells per trial 
        # (n will increment, to test scalability of performance)
        subset <- 
            sobj[, sample(x = 1:ncol(sobj), size = subset_size, replace = FALSE)]
        
        # Form filename for subset
        subset_filename <- paste0(subset_size, "_", i, ".rds")
        
        # Save subset to disk
        saveRDS(
            subset, 
            file = 
                paste0(
                  working_dir,
                  "Random_subsets/Seurat/",
                  subset_filename
                )
            )
        cat("\r")
    }
}

```

SingleCellExperiment objects

```{r}
# Generate SCE object from each .Rds object from above code block
for (subset_size in subset_sizes){
    for (i in trials){ 
        gc()
        rds_path <- 
            paste0(
              working_dir,
              "Random_subsets/Seurat/",
              # as.integer to output 500000 instead of 5e+05
              paste0(as.integer(subset_size), "_", i, ".rds")
            )
        
        cat(
            subset_size, 
            " cells, ", 
            "trial ", 
            i, 
            " load Seurat object ", 
            rds_path, 
            sep = ""
            )
         cat("\r")
        
        # Load Seurat object for the ith trial of the current subset size,
        # and convert to an SCE object
        sobj <- 
            readRDS(
                rds_path
            )
        
        cat(subset_size, " cells, ", "trial ", i, " make SCE object", sep = "")
        cat("\r")
        
        # Only one assay present
        sce <- as.SingleCellExperiment(sobj, assay = "RNA")
        
        sce_path <- 
            paste0(
              working_dir,
              "Random_subsets/SingleCellExperiment/",
              paste0(as.integer(subset_size), "_", i)
            )
        
        cat(
            subset_size, 
            " cells, ", 
            "trial ", 
            i, 
            " save SCE object ", 
            sce_path, 
            sep = ""
            )
    
        # Save in HDF5 format (use subset size and trial number)
        saveHDF5SummarizedExperiment(
          sce, 
          dir = sce_path,
          level = 3,
          replace = TRUE
          )
        
        cat("\r")
        }
    }



```

Creation of anndata objects 

```{r}
#subset_sizes <- c(1000, 2000, 5000, 10000, 20000, 50000, 100000, 200000, 500000) |> as.integer()
#subset_sizes <- 500000
#trials <- 1:5

# Generate SCE object from each .Rds object from above code block
for (subset_size in subset_sizes){
    for (i in trials){ 
        rds_path <- 
            paste0(
              working_dir,
              "Random_subsets/Seurat/",
              # as.integer to output 500000 instead of 5e+05
              paste0(as.integer(subset_size), "_", i, ".rds")
            )
        
        cat(
            subset_size, 
            " cells, ", 
            "trial ", 
            i, 
            " load Seurat object ", 
            rds_path, 
            sep = ""
            )
         cat("\r")
        
        # Load Seurat object for the ith trial of the current subset size,
        # and convert to an SCE object
        sobj <-
          readRDS(
            rds_path
            )
        
        adata_path <- 
          paste0(
            working_dir,
            "Random_subsets/anndata/",
            paste0(as.integer(subset_size), "_", i, ".h5ad")
            )
        
        cat(
          subset_size, 
          " cells, ", 
          "trial ", 
          i, 
          " save Anndata object ", 
          adata_path, 
          sep = ""
          )
    
        # Convert to anndata object and save
        adata <- 
          sceasy::convertFormat(
            sobj, 
            from = "seurat", 
            to = "anndata",
            drop_single_values = FALSE
            )
        
        anndata::write_h5ad(
          adata, 
          filename = adata_path
          )
        
        cat("\r")
        }
    }




```

# Performance testing of FetchData on subsets

```{r}
times <- c()
n_cells <- c()

# Seurat objects
for (subset_size in subset_sizes){
    # Run test for i trials per number of cells
    for (i in trials){
        # Display the current subset size and trial
        cat("Subset size ", subset_size, ", trial ", i, sep = "")
        
        # Load random subset
        subset <- 
          readRDS(
            paste0(
              working_dir,
              "Random_subsets/Seurat/",
              paste0(as.integer(subset_size), "_", i, ".rds")
            )
          )
        
        # Run FetchData and time run with Sys.time()
        start_time <- Sys.time()
        
        FetchData(
            subset, 
            # CellXGene displays feature names according to Ensembl ID.
            # To avoid the hassle of transforming cell IDs to gene symbols, 
            # I put a random gene symbol in GeneCards and used the corresponding
            # ensembl ID as the target var.
            vars = c("ENSG00000078399")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        # Also store the number of cells associated with the current trial
        n_cells <- c(n_cells, subset_size)
        # Carriage return, to avoid cluttering output during testing
        cat("\r")
    }
}

seurat_trial <-
    data.frame(
    `n_cells` = n_cells,
    `run_time` = times
    )

seurat_trial$fxn <- "FetchData.Seurat"
```

```{r}
# FetchData, SingleCellExperiment
times <- c()
n_cells <- c()

for (subset_size in subset_sizes){
    # Run test for i trials per number of cells
    for (i in trials){
        # Display the current subset size and trial
        cat("Subset size ", subset_size, ", trial ", i, sep = "")
        
        # Load random subset
        subset <- 
          HDF5Array::loadHDF5SummarizedExperiment(
            paste0(
              working_dir,
              "Random_subsets/SingleCellExperiment/",
              paste0(as.integer(subset_size), "_", i)
              )
            )
        
        # Run FetchData and time run with Sys.time()
        start_time <- Sys.time()
        
        FetchData(
            subset, 
            vars = c("ENSG00000078399")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        # Also store the number of cells associated with the current trial
        n_cells <- c(n_cells, subset_size)
        # Carriage return, to avoid cluttering output during testing
        cat("\r")
    }
}

sce_trial <-
  data.frame(
    `n_cells` = n_cells,
    `run_time` = times
  )

sce_trial$fxn <- "FetchData.SingleCellExperiment"
```

```{r}
# FetchData, anndata objects
times <- c()
n_cells <- c()

for (subset_size in subset_sizes){
    # Run test for i trials per number of cells
    for (i in trials){
        # Display the current subset size and trial
        cat("Subset size ", subset_size, ", trial ", i, sep = "")
        
        # Load random subset
        subset <- 
          anndata::read_h5ad(
            paste0(
              working_dir,
              "Random_subsets/anndata/",
              paste0(as.integer(subset_size), "_", i, ".h5ad")
            )
          )
        
        # Run FetchData and time run with Sys.time()
        start_time <- Sys.time()
        
        FetchData(
            subset, 
            vars = c("ENSG00000078399")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        # Also store the number of cells associated with the current trial
        n_cells <- c(n_cells, subset_size)
        # Carriage return, to avoid cluttering output during testing
        cat("\r")
        
        rm(subset)
        gc()
    }
}

anndata_trial <-
    data.frame(
    `n_cells` = n_cells,
    `run_time` = times
    )

anndata_trial$fxn <- "FetchData.AnnDataR6"

```

```{r}
# Combine results for each object and save
fetchdata_results <- rbind(seurat_trial, sce_trial, anndata_trial)

results_dir <- paste0(working_dir, "performance_results/")

if(!dir.exists(results_dir)) dir.create(results_dir)

write.csv(
  fetchdata_results, 
  file = paste0(results_dir, "FetchData.csv"), 
  row.names = FALSE
)
```

### Plot

```{r fetchData plot}
results_df <- read.csv(paste0(working_dir, "performance_results/FetchData.csv"))

# Rename function column lavels
results_df$fxn <- 
  plyr::revalue(
    results_df$fxn, 
    c("FetchData.Seurat" = "Seurat", 
      "FetchData.SingleCellExperiment" = "SingleCellExperiment", 
      "FetchData.AnnDataR6" = "anndata"
      )
    )

# Re-factor
results_df$fxn <- 
  factor(
    results_df$fxn, 
    levels = 
      c("Seurat",
        "SingleCellExperiment",
        "anndata"
        )
    )

# Convert number of cells to character
results_df$n_cells <-
  as.character(
    # Extract from scientific notation with as.integer
    as.integer(results_df$n_cells)
    )

results_df$n_cells <- 
  factor(
    results_df$n_cells,
    levels = 
      as.character(
        str_sort(
          unique(results_df$n_cells), 
          numeric = TRUE
          )
        )
  )

results_df <- results_df %>% rename(c("Method" = "fxn"))

fetchData_comparison <- 
  ggplot(data = results_df, aes(x = n_cells, y = run_time)) +
  geom_boxplot(aes(color = Method)) +
  scale_color_manual(
    values = 
      c("#880000", "#000088", "#B8AB2F"),
    aesthetics = "color",
    na.value = "grey50"
    ) +
  # labs(
  #   title = "FetchData with a Single Gene", 
  #   #subtitle = "Assessed over 100 trials, on full AML dataset, with one feature"
  #   ) +
  xlab("Number of cells") +
  ylab("Run Time (s)") +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    plot.subtitle = element_text(hjust = 0.5),
    #axis.title.y = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) #+
  #coord_flip()

fetchData_comparison
```

```{r save plot}
ggsave(
  fetchData_comparison,
  filename = paste0(working_dir,"performance_results/FetchData.png"), 
  device = "png", 
  width = 16, 
  height = 8,
  bg = "#FFFFFF"
  )
```


## fetchData on one metadata variable vs. fetch_metadata

```{r}
#subset_sizes <- c(1000, 2000, 5000, 10000, 20000, 50000, 100000, 200000)
#trials <- 1:5

times <- c()
n_cells <- c()
fxn <- c()

# Seurat objects
for (subset_size in subset_sizes){
    # Run test for i trials per number of cells
    for (i in trials){
        # Display the current subset size and trial
        cat("Subset size ", subset_size, ", trial ", i, sep = "")
        
        # Load random subset
        subset <- 
          readRDS(
            paste0(
              working_dir, 
              "Random_subsets/Seurat/",
              paste0(as.integer(subset_size), "_", i, ".rds")
            )
          )
        
        # Run FetchData and time run with Sys.time()
        start_time <- Sys.time()
        
        FetchData(
            subset, 
            vars = c("cell_type")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        # Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "FetchData.Seurat")
        
        start_time <- Sys.time()
        
        SCUBA::fetch_metadata(
            subset, 
            vars = c("cell_type")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        # Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "fetch_metadata.Seurat")
        
        # Carriage return, to avoid cluttering output during testing
        cat("\r")
    }
}

seurat_trial <-
    data.frame(
    `n_cells` = n_cells,
    `run_time` = times,
    `fxn` = fxn
    )
```

```{r}
# FetchData, SingleCellExperiment
#subset_sizes <- c(1000, 2000, 5000, 10000, 20000, 50000, 100000, 200000)
#trials <- 1:5

times <- c()
n_cells <- c()
fxn <- c()

for (subset_size in subset_sizes){
    # Run test for i trials per number of cells
    for (i in trials){
        # Display the current subset size and trial
        cat("Subset size ", subset_size, ", trial ", i, sep = "")
        
        # Load random subset
        subset <- 
          HDF5Array::loadHDF5SummarizedExperiment(
            paste0(
              working_dir,
              "Random_subsets/SingleCellExperiment/",
              paste0(as.integer(subset_size), "_", i)
              )
            )
        
        # Run FetchData and time run with Sys.time()
        start_time <- Sys.time()
        
        FetchData(
            subset, 
            vars = c("cell_type")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        #  Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "FetchData.SingleCellExperiment")
        
        start_time <- Sys.time()
        
        SCUBA::fetch_metadata(
            subset, 
            vars = c("cell_type")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        # Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "fetch_metadata.SingleCellExperiment")
        # Carriage return, to avoid cluttering output during testing
        cat("\r")
    }
}

sce_trial <-
  data.frame(
    `n_cells` = n_cells,
    `run_time` = times,
    `fxn` = fxn
  )
```

```{r}
#subset_sizes <- c(1000, 2000, 5000, 10000, 20000, 50000, 100000, 200000)
#trials <- 1:5

# FetchData, anndata objects
times <- c()
n_cells <- c()
fxn <- c()

# Seurat objects
for (subset_size in subset_sizes){
    # Run test for i trials per number of cells
    for (i in trials){
        # Display the current subset size and trial
        cat("Subset size ", subset_size, ", trial ", i, sep = "")
        
        # Load random subset
        subset <- 
          anndata::read_h5ad(
            paste0(
              working_dir,
              "Random_subsets/anndata/",
              paste0(as.integer(subset_size), "_", i, ".h5ad")
            )
          )
        
        # Run FetchData and time run with Sys.time()
        start_time <- Sys.time()
        
        FetchData(
            subset, 
            vars = c("cell_type")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        #  Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "FetchData.AnnDataR6")
        
        start_time <- Sys.time()
        
        SCUBA::fetch_metadata(
            subset, 
            vars = c("cell_type")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        
        # Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "fetch_metadata.AnnDataR6")
        
        # Carriage return, to avoid cluttering output during testing
        cat("\r")
    }
}

anndata_trial <-
    data.frame(
    `n_cells` = n_cells,
    `run_time` = times,
    `fxn` = fxn
    )

#have to run this after big anndata chunks
reticulate::import("gc")$collect()

```

```{r}
fetchdata_vs_fetchmetadata_results <- 
  rbind(
    seurat_trial, 
    sce_trial, 
    anndata_trial
  )

write.csv(
  fetchdata_vs_fetchmetadata_results,
  file = paste0(
    working_dir,
    "performance_results/FetchData_vs_fetch_metadata_3.csv"
  ),
  row.names = FALSE
)
```

```{r FetchData vs. fetch_metadata drafts}
# # Boxplot ####
# results_df <- read.csv("SCUBA/performance_results/FetchData_vs_fetch_metadata_3.csv")
# 
# # Add information for object type and function based on `fxn` column
# results_df$object <- 
#   case_when(
#     results_df$fxn %in% 
#       c("FetchData.Seurat", "fetch_metadata.Seurat") ~ "Seurat",
#     results_df$fxn %in% 
#       c("FetchData.SingleCellExperiment", 
#         "fetch_metadata.SingleCellExperiment") ~ "SingleCellExperiment",
#     results_df$fxn %in%
#       c("FetchData.AnnDataR6", "fetch_metadata.AnnDataR6") ~ "anndata",
#     TRUE ~ NA
#   )
# 
# results_df$Function <- 
#   case_when(
#     results_df$fxn %in% 
#       c("FetchData.Seurat", 
#         "FetchData.SingleCellExperiment", 
#         "FetchData.AnnDataR6") ~ "FetchData",
#     results_df$fxn %in% 
#       c("fetch_metadata.Seurat",  
#         "fetch_metadata.SingleCellExperiment",
#         "fetch_metadata.AnnDataR6") ~ "fetch_metadata",
#     TRUE ~ NA
#   )
# 
# # Re-factor object, function columns
# results_df$object <-
#   factor(
#     results_df$object,
#     levels = c("Seurat", "SingleCellExperiment", "anndata")
#   )
# 
# results_df$Function <-
#   factor(
#     results_df$Function,
#     levels = c("FetchData", "fetch_metadata")
#   )
# 
# 
# # Convert number of cells to character
# results_df$n_cells <-
#   as.character(
#     # Extract from scientific notation with as.integer
#     as.integer(results_df$n_cells)
#     )
# 
# results_df$n_cells <- 
#   factor(
#     results_df$n_cells,
#     levels = 
#       str_sort(
#         unique(results_df$n_cells), 
#         numeric = TRUE
#         )
#   )
# 
# # Remove single outlier for SingleCellExperiment, FetchData, 10,000 cells
# results_df <- results_df[-113,] 
# 
# #results_df <- results_df %>% rename(c("Method" = "fxn"))
# 
# fetchData_box_plt <- 
#   ggplot(data = results_df, aes(x = object, y = run_time)) +
#   geom_boxplot(aes(color = Function)) +
#   facet_wrap(
#     ~ n_cells, 
#     #scale = "free", 
#     ncol = 4
#     ) +
#   scale_color_manual(
#     values = 
#       c("#0079FF", "#8C00FF"), # #yellow "#F2AF0D", #orange "#F25E0D", #green "#59A662", #Red "#E61A1A"
#     aesthetics = "color",
#     na.value = "grey50"
#     ) +
#   labs(
#     title = "FetchData vs. fetch_metadata", 
#     #subtitle = "Assessed over 100 trials, on full AML dataset, with one feature"
#     ) +
#   #xlab("Number of cells") +
#   ylab("Run Time (s)") +
#   theme(
#     plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
#     plot.subtitle = element_text(hjust = 0.5),
#     #axis.title.y = element_blank(),
#     axis.title.x = element_blank(),
#     axis.text.x = element_text(angle = 45, hjust = 1)
#   ) +
#   coord_cartesian(ylim = c(0,0.5)) +
#   facet_theme #+
#   #coord_flip()
# 
# fetchData_box_plt
# 
# ggsave(
#   fetchData_box_plt,
#   filename = "SCUBA/performance_results/FetchData_vs_FetchMetadata_3_fixed_scale.png",
#   device = "png",
#   width = 16,
#   height = 8,
#   bg = "#FFFFFF"
#   )
# 
# # Boxplot, separating object types, focusing on comparison between functions
# fetchdata_box_plots <- list()
# 
# for (obj_type in c("Seurat", "SingleCellExperiment", "anndata")){
#   fetchdata_box_plots[[obj_type]] <-
#   results_df %>% 
#   dplyr::filter(object == obj_type) %>% 
#   ggplot(data = ., aes(x = n_cells, y = run_time)) +
#   geom_boxplot(aes(color = Function)) +
#   # facet_wrap(
#   #   ~ n_cells, 
#   #   #scale = "free", 
#   #   ncol = 4
#   #   ) +
#   scale_color_manual(
#     values = 
#       c("#0079FF", "#8C00FF"), # #yellow "#F2AF0D", #orange "#F25E0D", #green "#59A662", #Red "#E61A1A"
#     aesthetics = "color",
#     na.value = "grey50"
#     ) +
#   labs(
#     title = "FetchData vs. fetch_metadata", 
#     #subtitle = "Assessed over 100 trials, on full AML dataset, with one feature"
#     ) +
#   #xlab("Number of cells") +
#   ylab("Run Time (s)") +
#   theme(
#     plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
#     plot.subtitle = element_text(hjust = 0.5),
#     #axis.title.y = element_blank(),
#     axis.title.x = element_blank(),
#     axis.text.x = element_text(angle = 45, hjust = 1)
#   ) +
#   coord_cartesian(ylim = c(0,0.5)) +
#   facet_theme #+
#   #coord_flip()
# }
# 
# # Plots split by object type, separating scales for each object ####
# obj_types <- c("Seurat", "SingleCellExperiment", "anndata")
# fetchdata_vs_fetchmetadata_box <- list()
# 
# for (i in 1:length(obj_types)){
#   fetchdata_vs_fetchmetadata_box[[i]] <-
#     results_df %>% 
#     dplyr::filter(object == obj_types[[i]]) %>% 
#     ggplot(data = ., aes(x = n_cells, y = run_time)) +
#     geom_boxplot(aes(color = Function)) +
#     # facet_wrap(
#     #   ~ n_cells, 
#     #   #scale = "free", 
#     #   ncol = 4
#     #   ) +
#     scale_color_manual(
#       values = 
#         c("#0079FF", "#8C00FF"), # #yellow "#F2AF0D", #orange "#F25E0D", #green "#59A662", #Red "#E61A1A"
#       aesthetics = "color",
#       na.value = "grey50"
#       ) +
#     # labs(
#     #   title = "FetchData vs. fetch_metadata", 
#     #   #subtitle = "Assessed over 100 trials, on full AML dataset, with one feature"
#     #   ) +
#     xlab("Number of Cells") +
#     ylab("Run Time (s)") +
#     theme(
#       plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
#       plot.subtitle = element_text(hjust = 0.5),
#       #axis.title.y = element_blank(),
#       axis.title.x = element_blank(),
#       axis.text.x = element_text(angle = 45, hjust = 1)
#     ) +
#     #coord_cartesian(ylim = c(0,0.5)) +
#     facet_theme #+
#     #coord_flip()
# }
# 
# patchwork::wrap_plots(
#   fetchdata_vs_fetchmetadata_box,
#   ncol = 1,
#   guides = "collect"
# )
# 
# 
# # Alternate: facet by object instead of number of cells ####
# facet_by_object <-
#   results_df %>%
#   ggplot(data = ., aes(x = n_cells, y = run_time)) +
#   geom_boxplot(aes(color = Function)) +
#   facet_wrap(
#     ~ object,
#     scales = "free",
#     ncol = 1
#     ) +
#   scale_color_manual(
#     values = 
#       c("#0079FF", "#8C00FF"), # #yellow "#F2AF0D", #orange "#F25E0D", #green "#59A662", #Red "#E61A1A"
#     aesthetics = "color",
#     na.value = "grey50"
#     ) +
#   labs(
#     title = "FetchData vs. fetch_metadata", 
#     #subtitle = "Assessed over 100 trials, on full AML dataset, with one feature"
#     ) +
#   ylab("Run Time (s)") +
#   xlab("Number of Cells") +
#   theme(
#     plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
#     plot.subtitle = element_text(hjust = 0.5),
#     #axis.title.y = element_blank(),
#     #axis.title.x = element_blank(),
#     axis.text.x = element_text(angle = 45, hjust = 1)
#   ) +
#   #coord_cartesian(ylim = c(0,0.5)) +
#   facet_theme
#  
# facet_by_object

```

```{r fetchData_vs_fetch_metadata_boxplot}
# Boxplot ####
results_df <- read.csv(paste0(working_dir, "performance_results/FetchData_vs_fetch_metadata_3.csv"))

# Add information for object type and function based on `fxn` column
results_df$object <- 
  case_when(
    results_df$fxn %in% 
      c("FetchData.Seurat", "fetch_metadata.Seurat") ~ "Seurat",
    results_df$fxn %in% 
      c("FetchData.SingleCellExperiment", 
        "fetch_metadata.SingleCellExperiment") ~ "SingleCellExperiment",
    results_df$fxn %in%
      c("FetchData.AnnDataR6", "fetch_metadata.AnnDataR6") ~ "anndata",
    TRUE ~ NA
  )

results_df$Function <- 
  case_when(
    results_df$fxn %in% 
      c("FetchData.Seurat", 
        "FetchData.SingleCellExperiment", 
        "FetchData.AnnDataR6") ~ "FetchData",
    results_df$fxn %in% 
      c("fetch_metadata.Seurat",  
        "fetch_metadata.SingleCellExperiment",
        "fetch_metadata.AnnDataR6") ~ "fetch_metadata",
    TRUE ~ NA
  )

# Re-factor object, function columns
results_df$object <-
  factor(
    results_df$object,
    levels = c("Seurat", "SingleCellExperiment", "anndata")
  )

results_df$Function <-
  factor(
    results_df$Function,
    levels = c("FetchData", "fetch_metadata")
  )


# Convert number of cells to character
results_df$n_cells <-
  as.character(
    # Extract from scientific notation with as.integer
    as.integer(results_df$n_cells)
    )

results_df$n_cells <- 
  factor(
    results_df$n_cells,
    levels = 
      str_sort(
        unique(results_df$n_cells), 
        numeric = TRUE
        )
  )

# Remove single outlier for SingleCellExperiment, FetchData, 10,000 cells
results_df <- results_df[-113,] 

# Plot
fetchData_vs_fetch_metadata <-
  results_df %>%
  ggplot(data = ., aes(x = n_cells, y = run_time)) +
  geom_boxplot(aes(color = Function)) +
  facet_wrap(
    ~ object,
    scales = "free",
    ncol = 1
    ) +
  scale_color_manual(
    values = 
      c("#0079FF", "#8C00FF"), # #yellow "#F2AF0D", #orange "#F25E0D", #green "#59A662", #Red "#E61A1A"
    aesthetics = "color",
    na.value = "grey50"
    ) +
  # labs(
  #   title = "FetchData vs. fetch_metadata", 
  #   #subtitle = "Assessed over 100 trials, on full AML dataset, with one feature"
  #   ) +
  ylab("Run Time (s)") +
  xlab("Number of Cells") +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    plot.subtitle = element_text(hjust = 0.5),
    #axis.title.y = element_blank(),
    #axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  #coord_cartesian(ylim = c(0,0.5)) +
  facet_theme
 
fetchData_vs_fetch_metadata
```

```{r}
# Save box plot from above code chunk (version faceted by object)
ggsave(
  fetchData_vs_fetch_metadata,
  filename = paste0(working_dir, "performance_results/FetchData_vs_FetchMetadata_3_free_scale_facet_by_object.png"),
  device = "png",
  width = 16,
  height = 8,
  bg = "#FFFFFF"
  )
```

```{r FetchData vs. fetch_metadata lineplot}
# results_df <- read.csv("SCUBA/performance_results/FetchData_vs_fetch_metadata_3.csv")
# 
# # Add information for object type and function based on `fxn` column
# results_df$object <- 
#   case_when(
#     results_df$fxn %in% 
#       c("FetchData.Seurat", "fetch_metadata.Seurat") ~ "Seurat",
#     results_df$fxn %in% 
#       c("FetchData.SingleCellExperiment", 
#         "fetch_metadata.SingleCellExperiment") ~ "SingleCellExperiment",
#     results_df$fxn %in%
#       c("FetchData.AnnDataR6", "fetch_metadata.AnnDataR6") ~ "anndata",
#     TRUE ~ NA
#   )
# 
# results_df$Function <- 
#   case_when(
#     results_df$fxn %in% 
#       c("FetchData.Seurat", 
#         "FetchData.SingleCellExperiment", 
#         "FetchData.AnnDataR6") ~ "FetchData",
#     results_df$fxn %in% 
#       c("fetch_metadata.Seurat",  
#         "fetch_metadata.SingleCellExperiment",
#         "fetch_metadata.AnnDataR6") ~ "fetch_metadata",
#     TRUE ~ NA
#   )
# 
# # Re-factor object, function columns
# results_df$object <-
#   factor(
#     results_df$object,
#     levels = c("Seurat", "SingleCellExperiment", "anndata")
#   )
# 
# results_df$Function <-
#   factor(
#     results_df$Function,
#     levels = c("FetchData", "fetch_metadata")
#   )
# 
# 
# # Convert number of cells to character
# results_df$n_cells <-
#   as.character(
#     # Extract from scientific notation with as.integer
#     as.integer(results_df$n_cells)
#     )
# 
# results_df$n_cells <- 
#   factor(
#     results_df$n_cells,
#     levels = 
#       str_sort(
#         unique(results_df$n_cells), 
#         numeric = TRUE
#         )
#   )
# 
# # Remove single outlier for SingleCellExperiment, FetchData, 10,000 cells
# results_df <- results_df[-113,] 
# 
# # Create summary stats for results
# summary_df <-
#   results_df |>
#   group_by(Function, object, n_cells) |>
#   summarize(
#     mean_time = mean(run_time),
#     sd = sd(run_time)
#   )
# 
# line_plot <-
#   ggplot(data = summary_df, aes(x = n_cells, y = mean_time, color = Function, group = Function)) +
#   geom_point() +
#   geom_line() +
#   geom_errorbar(aes(ymin = mean_time - sd, ymax = mean_time + sd)) +
#   scale_color_manual(
#       values = c("#0079FF", "#8C00FF"),
#       aesthetics = "color",
#       na.value = "grey50"
#       ) +
#   facet_wrap(
#     ~ object,
#     ncol = 1,
#     scales = "free"
#   ) +
#   # scale_y_continuous(
#   #   limits = c(0, 9)
#   # ) +
#   # scale_y_continuous(
#   #   breaks = seq(from = 0, to = 0.5, by = 0.05),
#   #   limits = c(0, 0.5)#,
#   #   #labels = function(breaks){paste0(breaks, " s")}
#   #   ) +
#   ylab("Time (s)") +
#   xlab("Number of Cells") +
#   labs(
#     title = "FetchData vs. fetch_metadata",
#     # subtitle =
#     #   "(200k Cells x 29k Genes, HDF5Array Compression Level 3 for SCE objects)"
#   ) +
#   guides(
#     color =
#       guide_legend(
#         title = NULL
#         )
#   ) +
#   theme_cowplot() +
#   theme(
#     plot.title = element_text(hjust = 0.5, size = 18),
#     plot.subtitle = element_text(hjust = 0.5),
#     axis.text.x = element_text(size = 9)
#     ) +
#   facet_theme
# 
# line_plot
```

```{r}
# Save line plot
# ggsave(
#   line_plot,
#   filename = "SCUBA/performance_results/FetchData_vs_FetchMetadata_3_line_plot.png",
#   device = "png",
#   width = 16,
#   height = 8,
#   bg = "#FFFFFF"
#   )
```

## fetchData on UMAP coordinates vs. fetch_reduction

```{r}
#subset_sizes <- c(1000, 2000, 5000, 10000, 20000, 50000, 100000, 200000)
#trials <- 1:5

times <- c()
n_cells <- c()
fxn <- c()

# Seurat objects
for (subset_size in subset_sizes){
    # Run test for i trials per number of cells
    for (i in trials){
        # Display the current subset size and trial
        cat("Subset size ", subset_size, ", trial ", i, sep = "")
        
        # Load random subset
        subset <- 
          readRDS(
            paste0(
              working_dir,
              "Random_subsets/Seurat/",
              paste0(as.integer(subset_size), "_", i, ".rds")
            )
          )
        
        # Run FetchData and time run with Sys.time()
        start_time <- Sys.time()
        
        FetchData(
            subset, 
            vars = c("UMAP_1", "UMAP_2")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        # Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "FetchData.Seurat")
        
        start_time <- Sys.time()
        
        SCUBA::fetch_reduction(
            subset, 
            reduction = "UMAP",
            dims = c(1, 2)
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        # Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "fetch_reduction.Seurat")
        
        # Carriage return, to avoid cluttering output during testing
        cat("\r")
    }
}

seurat_trial <-
    data.frame(
    `n_cells` = n_cells,
    `run_time` = times,
    `fxn` = fxn
    )
```

```{r}
# FetchData, SingleCellExperiment
#subset_sizes <- c(1000, 2000, 5000, 10000, 20000, 50000, 100000, 200000)
#trials <- 1:5

times <- c()
n_cells <- c()
fxn <- c()

for (subset_size in subset_sizes){
    # Run test for i trials per number of cells
    for (i in trials){
        # Display the current subset size and trial
        cat("Subset size ", subset_size, ", trial ", i, sep = "")
        
        # Load random subset
        subset <- 
          HDF5Array::loadHDF5SummarizedExperiment(
            paste0(
              working_dir,
              "Random_subsets/SingleCellExperiment/",
              paste0(as.integer(subset_size), "_", i)
              )
            )
        
        # Run FetchData and time run with Sys.time()
        start_time <- Sys.time()
        
        FetchData(
            subset, 
            vars = c("UMAP_1", "UMAP_2")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        #  Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "FetchData.SingleCellExperiment")
        
        start_time <- Sys.time()
        
        SCUBA::fetch_reduction( 
            subset, 
            reduction = "UMAP",
            dims = c(1, 2)
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        # Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "fetch_reduction.SingleCellExperiment")
        # Carriage return, to avoid cluttering output during testing
        cat("\r")
    }
}

sce_trial <-
  data.frame(
    `n_cells` = n_cells,
    `run_time` = times,
    `fxn` = fxn
  )
```



```{r}
#subset_sizes <- c(1000, 2000, 5000, 10000, 20000, 50000, 100000, 200000)
#trials <- 1:5

# FetchData, anndata objects
times <- c()
n_cells <- c()
fxn <- c()

# Seurat objects
for (subset_size in subset_sizes){
    # Run test for i trials per number of cells
    for (i in trials){
        # Display the current subset size and trial
        cat("Subset size ", subset_size, ", trial ", i, sep = "")
        
        # Load random subset
        subset <- 
          anndata::read_h5ad(
            paste0(
              working_dir,
              "Random_subsets/anndata/",
              paste0(as.integer(subset_size), "_", i, ".h5ad")
            )
          )
        
        # Run FetchData and time run with Sys.time()
        start_time <- Sys.time()
        
        FetchData(
            subset, 
            vars = c("X_umap_1", "X_umap_2")
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        #  Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "FetchData.AnnDataR6")
        
        start_time <- Sys.time()
        
        SCUBA::fetch_reduction(
            subset, 
            reduction = "X_umap",
            dims = c(1, 2)
            )
        
        end_time <- Sys.time()
        
        # Compute time elapsed with difftime and append to vector of times
        time_i <- as.numeric(difftime(end_time, start_time, units = 'secs'))
        
        times <- c(times, time_i)
        
        # Also store the number of cells associated with the current trial,
        # and the function ran 
        n_cells <- c(n_cells, subset_size)
        fxn <- c(fxn, "fetch_reduction.AnnDataR6")
        
        # Carriage return, to avoid cluttering output during testing
        cat("\r")
    }
}

anndata_trial <-
    data.frame(
    `n_cells` = n_cells,
    `run_time` = times,
    `fxn` = fxn
    )

#required after large anndata chunks
reticulate::import("gc")$collect()
```

```{r}
fetchdata_vs_fetchreduction_results <- 
  rbind(seurat_trial, sce_trial, anndata_trial)

write.csv(
   fetchdata_vs_fetchreduction_results,
   file = paste0(
     working_dir,
     "performance_results/FetchData_vs_fetch_reduction.csv"
   ),
   row.names = FALSE
)
```

## Box plot

```{r FetchData vs. fetch_reduction boxplot}
results_df <- 
  read.csv(
    paste0(
      working_dir,
      "performance_results/FetchData_vs_fetch_reduction.csv"
    )
)

# Add information for object type and function based on `fxn` column
results_df$object <- 
  case_when(
    results_df$fxn %in% 
      c("FetchData.Seurat", "fetch_reduction.Seurat") ~ "Seurat",
    results_df$fxn %in% 
      c("FetchData.SingleCellExperiment", 
        "fetch_reduction.SingleCellExperiment") ~ "SingleCellExperiment",
    results_df$fxn %in%
      c("FetchData.AnnDataR6", "fetch_reduction.AnnDataR6") ~ "anndata",
    TRUE ~ NA
  )

results_df$Function <- 
  case_when(
    results_df$fxn %in% 
      c("FetchData.Seurat", 
        "FetchData.SingleCellExperiment", 
        "FetchData.AnnDataR6") ~ "FetchData",
    results_df$fxn %in% 
      c("fetch_reduction.Seurat",  
        "fetch_reduction.SingleCellExperiment",
        "fetch_reduction.AnnDataR6") ~ "fetch_reduction",
    TRUE ~ NA
  )

# Re-factor object, function columns
results_df$object <-
  factor(
    results_df$object,
    levels = c("Seurat", "SingleCellExperiment", "anndata")
  )

results_df$Function <-
  factor(
    results_df$Function,
    levels = c("FetchData", "fetch_reduction")
  )


# Convert number of cells to character
results_df$n_cells <-
  as.character(
    # Extract from scientific notation with as.integer
    as.integer(results_df$n_cells)
    )

results_df$n_cells <- 
  factor(
    results_df$n_cells,
    levels = 
      str_sort(
        unique(results_df$n_cells), 
        numeric = TRUE
        )
  )

# Remove extreme outlier for SingleCellExperiment, fetch_reduction
results_df <- results_df[-93,]

fetchdata_vs_fetch_reduction <-
  results_df %>%
  ggplot(data = ., aes(x = n_cells, y = run_time)) +
  geom_boxplot(aes(color = Function)) +
  facet_wrap(
    ~ object,
    scales = "free",
    ncol = 1
    ) +
  scale_color_manual(
    values = 
      c("#0079FF", "#E61A1A"), # #yellow "#F2AF0D", #orange "#F25E0D", #green "#59A662", #Red "#E61A1A"
    aesthetics = "color",
    na.value = "grey50"
    ) +
  # labs(
  #   title = "FetchData vs. fetch_metadata", 
  #   #subtitle = "Assessed over 100 trials, on full AML dataset, with one feature"
  #   ) +
  ylab("Run Time (s)") +
  xlab("Number of Cells") +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    plot.subtitle = element_text(hjust = 0.5),
    #axis.title.y = element_blank(),
    #axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  #coord_cartesian(ylim = c(0,0.5)) +
  facet_theme
 
fetchdata_vs_fetch_reduction
```

```{r}
# Save box plot
ggsave(
  fetchdata_vs_fetch_reduction,
  filename = paste0(
    working_dir,
    "performance_results/FetchData_vs_fetch_reduction.png"
  ),
  device = "png",
  width = 16,
  height = 8,
  bg = "#FFFFFF"
)
```

# Combined figure

```{r}
layout <- 
  c("
    AAAA
    BBCC
    BBCC
    ")

combined_figure <-
  patchwork::wrap_plots(
    fetchData_comparison,
    fetchData_vs_fetch_metadata,
    fetchdata_vs_fetch_reduction,
    design = layout,
    guides = "keep"
    ) +
    patchwork::plot_annotation(
        tag_levels = list(c("A", "B", "C"))
      ) &
  theme(plot.tag = element_text(size = 24, face = "bold"))

# PDF export
pdf(file = paste0(working_dir, "performance_results/combined_performance.pdf"), width = 14, height = 16)
combined_figure
dev.off()

# SVG Export
svglite::svglite(file = paste0(working_dir, "performance_results/combined_performance.svg"), width = 14, height = 16)
combined_figure
dev.off()

# PNG export
png(
  file = paste0(working_dir, 
                "performance_results/combined_performance.png"
         ), 
  width = 14, 
  height = 16, 
  units = "in", 
  res = 300
  )
combined_figure
dev.off()
```

