---
title: "Figure 6"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
library(Seurat)
library(SCUBA)
library(magrittr)
library(cowplot)
library(ggplot2)
library(dplyr)

library(SingleCellExperiment)
library(reticulate)
library(anndata)
```

The Rmd below contains code for generating **Figure 6** in the SCUBA manuscript. For instructions on generating the `AML_Seurat`, `AML_SCE`, and `AML_h5ad()` datasets, see "Demo_Object_Generation.Rmd".

## SCUBA simplifies common object exploration methods 

Below is a simple example operation using SCUBA: the identification of all unique values of a metadata variable represetned in the object. The process using SCUBA is compared to the most efficient equivalents for each object type.

### With SCUBA

```{r}
SCUBA::fetch_metadata(
  # May replace this with `AML_SCE()` or `AML_h5ad()` with no change in results
  AML_Seurat,
  # var is set to a metadata variable in the object.
  # To view available metadata, use SCUBA::meta_varnames().
  var = "condensed_cell_type"
  ) %>% 
  unique()
```

### Most efficient equivalents

Below are the most efficient equivalent operations for each object. Compared to SCUBA, these operations require the memorization of different functions for each object class, and require knowledge of each object structure to perform correctly.

```{r}
# Seurat objects
AML_Seurat$condensed_cell_type %>% 
  unique()

# SingleCellExperiment objects
colData(AML_SCE())$condensed_cell_type %>% 
  unique()

# anndata objects
AML_h5ad()$obs$condensed_cell_type %>% 
  unique()
```


